---
title: "GetCleanPBDB"
output:
  html_document: default
  word_document: default
  pdf_document: default
date: "2023-23-07"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


# setup environment
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(velociraptr,palaeoverse,deeptime,tidyverse,divDyn,ape,phytools,paleotree,dispRity,ggalt,dplyr,ggplot2,RCurl,janitor,fossilbrush)
```

# download the PBDB data using the source API
# All dinosaur occurances

```{r}
url <- "https://paleobiodb.org/data1.2/occs/list.csv?base_name=dinosauria&interval=Triassic,Cretaceous&show=full,classext,genus,acconly,ident,img,etbasis,class"

occs_df <- read.csv(url)
```


# now do some tidying up of the data by removing unwanted columns and remap any min_age that falls outside the Mesozoic.
# Also remove any occurance with no order, family or genus
# remove ootaxa and ichnotaxa


```{r}
# remove unwanted columns

keep <- c("accepted_name","accepted_rank","early_interval","late_interval","max_ma","min_ma","phylum","class","order","family","genus","lng","lat","cc","state","county","paleolng","paleolat","formation","stratgroup","member","assembl_comps")

occs_df <- occs_df[keep]

occs_df$min_ma[occs_df$min_ma < 66] <- 66
occs_df$max_ma[occs_df$max_ma > 251.902] <- 251.902

#now modify the ealy and later interval column to remove "Early, Middle or Late"
occs_df$early_interval <- sub('Early ','',occs_df$early_interval)
occs_df$early_interval <- sub('Middle ','',occs_df$early_interval)
occs_df$early_interval <- sub('Late ','',occs_df$early_interval)

occs_df$late_interval <- sub('Early ','',occs_df$late_interval)
occs_df$late_interval <- sub('Middle ','',occs_df$late_interval)
occs_df$late_interval <- sub('Late ','',occs_df$late_interval)

# use the fossilbrush package to clean up chronostratigraphic information to GTS2020 values

occs_df <- chrono_scale(occs_df, srt = "early_interval", end = "late_interval",max_ma = "max_ma", min_ma = "min_ma", verbose = FALSE)

# create time bins for Mesozoic Stages
MesozoicStage_bins <- time_bins(interval = c("Mesozoic"),rank="stage")

# bin the occurance data into new DF
MesozoicStage_Occs  <- bin_time(occs_df,MesozoicStage_bins) 
colnames(MesozoicStage_Occs)[colnames(MesozoicStage_Occs)=="bin_assignment"]="bin"
MesozoicStage_Occs <- merge(MesozoicStage_Occs,MesozoicStage_bins[,c("bin","interval_name","colour")],by="bin")


# rename the cols to fit later coding
# original min and max ma to PBDBmin_ma and PBDBmax_ma and the GTS2020 codes to GTS2020min_ma and GT2020max_ma
MesozoicStage_Occs <- MesozoicStage_Occs %>% 
  rename(
    min_maPBDB = min_ma,
    max_maPBDB = max_ma,
    GTS2020min_ma = newLAD,
    GTS2020max_ma = newFAD
    )


# extract microfossils

microverts <- 
  MesozoicStage_Occs %>%
  janitor::clean_names() %>%           # standardizes names
  filter(!grepl("oolithidae",family)) %>% #remove ootaxa and ichnotaxa
  filter(!grepl("podidae",family)) %>%
  filter(!grepl("podos",genus)) %>%
  filter(!grepl("podus",genus)) %>%
  filter(grepl("microfossils",assembl_comps)) %>% #extract microfossils
  filter(!grepl("Aves",class)) # remove Aves


MicrovertsByStage <- microverts %>%
  group_by(bin_midpoint) %>%
  summarise(count=n())



# extract macrofossils

macroverts <- 
  MesozoicStage_Occs %>%
  janitor::clean_names() %>%           # standardizes names
  filter(!grepl("oolithidae",family)) %>% #remove ootaxa and ichnotaxa
  filter(!grepl("podidae",family)) %>%
  filter(!grepl("podos",genus)) %>%
  filter(!grepl("podus",genus)) %>%
  filter(!grepl("microfossils",assembl_comps)) %>% #extract microfossils
  filter(!grepl("Aves",class)) # remove Aves


MacrovertsByStage <- macroverts %>%
  group_by(bin_midpoint) %>%
  summarise(count=n())

# combine the data

StageData <- merge(MicrovertsByStage,MacrovertsByStage,by="bin_midpoint")

StageData <- StageData %>% 
  rename(
    micro = count.x,
    macro = count.y
    )

StageData <- StageData %>% pivot_longer(., cols = c(micro, macro))

```

# create some plots

```{r}

p <- ggplot(StageData, aes(x=bin_midpoint, y=value)) +
  geom_bar(stat="identity",width=2.5, position="stack",aes(fill=name)) + 
  scale_fill_manual("",labels=c("Macro remains", "Microvertebrates"),values=c("#E69F00", "#0072B2")) + 
  scale_x_reverse() +
  ylab("Occurence count") + 
  xlab("") +
#  coord_geo(pos=as.list(rep("bottom",3)), dat=list("stages","epochs","periods"), height=list(unit(2,"lines"),unit(2,"lines"),unit(2,"line")), rot=list(90,0,0),size=list(2.5,2.5,5),abbrv=list(TRUE,FALSE,FALSE)) +
  coord_geo(pos=as.list(rep("bottom",2)), dat=list("epochs","periods"), height=list(unit(2,"lines"),unit(2,"lines")), rot=list(0,0),size=list(2.5,5),abbrv=list(FALSE,FALSE)) +
    theme_minimal() +
    theme(legend.position='top', 
        legend.justification='left',
        legend.direction='horizontal') 


pdf(file="PBDB_MIcrovertebrates.pdf", width=14/2.54, height=14/2.54)
p
dev.off()
